On the presentation of orthogonal groups over polynomial rings

### 2019-07-22 dl-paper.tex
Expanding preliminaries

\subsection{Unstable Quillen K-groups} \label{sec:quillen}
Throughout this section $G, P, S$ denote three group-valued functors on the category of commutative rings satisfying the following assumptions:
\begin{enumerate} [label=(A\arabic*)]
 \item \label{req:left-exact} $G$ is left exact;
 \item \label{req:subfunc} $P$ is a subfunctor of $G$ such that for all $R$ the group $P(R)$ is a perfect normal subgroup of $G(R)$;
 \item \label{req:uce} There is a natural transformation $\varphi \colon S \to P$ such that $ \varphi \colon S(R) \to P(R)$ is a universal central extension for all $R$;
 \item \label{req:coeq} For every pair $(R, I)$ the functor $S$ has the property that \[S(R \times_{R/I} R) \cong S(R) \times_{S(R/I)} S(R).\]
\end{enumerate}

@@

Let us also briefly recall Loday's procedure for relativization of $S$, that is,
 a way of extending $S$ from the category of rings to the category of pairs.
Denote by $p_0, p_1, \Delta$ the two obvious projections and the diagonal map $\xymatrix{D_{R,I} \ar@<1pt>[r] \ar@<-2.5pt>[r] & \ar@<-4pt>[l] R}$
  and set $d_i = Sp_i$, $s = S\Delta$.
 Obviously, $d_0s = d_1 s = id_{S(R)}$. Set $G_i = \Ker(d_i)$, and define $S(R, I)$ as $ G_0 / [G_0, G_1]$, denote by $\mu$ the natural map $S(R, I) \to S(R)$ induced by $d_1$.

%TODO: Give the definition of $D_{R, I}$ and of the category of Pairs
\begin{prop}\label{characterization}
For $(G, P, R)$ as above the following statements are true.
\begin{enumerate}
 \item There are natural isomorphisms
  \[ \K_2(G, P, R) \cong \Ker(S(R) \to G(R)), \ \K_3(G, P, R) \cong \HH_3(S(R), \ZZ). \]
 \item The map $\mu \colon S(R, I) \to S(R)$ is the universal relative central extension of the map $S(R) \to S(R/I)$.
       Denote the kernel of $\mu$ by $C(R, I)$.
       Then $C(R, I)$ is naturally isomorphic to the third relative homology group $\HH_3(S(R), S(R/I), \ZZ)$.
 \item There is a natural exact sequence:
 \[ \xymatrix{\K_3(G, P, R) \ar[r] & \K_3(G, P, R/I) \ar@{->>}[r] & C(R, I).} \]
 \end{enumerate}   
\end{prop}
\begin{proof}
 The first claim follows from~\ref{req:subfunc} and~\ref{req:uce} using the standard properties of the plus-construction, see \cite[\S~IV.1]{Kbook}   (cf. Exercises~1.8--1.9 ibid.). %provided one verifies that all the intermediate isomorphisms are natural in $R$.

>>>>

\subsection{Relative Steinberg groups} \label{sec:quillen}
In this subsection we recall the definitions and basic facts pertaining to relative Steinberg groups as defined by J.-L.~Loday in~\cite{Lo78},
 where a theory of relative central extensions is developed and then applied to the usual (stable) Steinberg group $\St(R)$ from algebraic K-theory (see e.\,g.~\cite[Ch. III \S~5]{Kbook}).
The main goal of this subsection is to show that Loday's result can be applied to the unstable Steinberg groups, and that
 the resulting relative groups enjoy the same good properties as their stable counterparts.

%Some of the results of this subsection have been already mentioned in~\cite{S15} (cf. e.\,g. Corollaries 3--4).
%Nevertheless, we chose to reproduce them with greater amount of detail for the reason that~\cref{lem:prop41} depends on the unstable exact sequence of.

First of all, let us recall the definition of the category of (commutative) pairs $\catname{Pairs}$.
The objects of $\catname{Pairs}$ are pairs $(R, I)$, in which $R$ is a commutative ring and $I$ its ideal.
By definition, a morphism $f \colon (R, I) \to (R', I')$ in the category of pairs corresponds to a ring map $f \colon R \to R'$ such that $f(I) \subseteq I'$.
Notice that the mapping $(R, I) \mapsto (R \to R/I)$ defines a functor from $\catname{Pairs}$ the morphism category $\catname{CRings}^\rightarrow$.

There is an obvious fully faithful embedding $\catname{CRings} \to \catname{Pairs}$ sending $R$ to $(R, R)$.
For a given functor $S \colon \catname{CRings} \to \catname{Groups}$ a {\it relativization} of $S$ is any functor $\widetilde{S} \colon \catname{Pairs} \to \catname{Groups}$
 extending $S$ in the obvious sense. Relativization of a functor is not unique.

Recall that {\it the double ring} $D_{R, I}$ of a pair $(R, I)$ is, by definition, the pullback ring $R \times_{R/I} R$.
In other words, it is the ring consisting of pairs of elements of $R$ congruent modulo $I$.
Denote by $p_0, p_1, \Delta$ the two obvious projections and the diagonal map $\xymatrix{D_{R,I} \ar@<1pt>[r] \ar@<-2.5pt>[r] & \ar@<-4pt>[l] R}$.

Let us describe the procedure for relativization as specified by Loday.
Let $S \colon \catname{CRings} \to \catname{Groups}$ be a functor.
Set $G_i = \Ker(S(p_i))$ and define Loday's relativization $S(R, I)$ as $ G_0 / [G_0, G_1]$. The map $S(p_1)$ induces a natural transformation $S(R, I) \to S(R)$. 
For the sake of shortness we denote this map by $\mu = \mu_{R,I}$. In fact, there is a natural exact sequence:
\begin{equation}
 \xymatrix{ 1 \ar[r] & C(R, I) \ar[r] & S(R, I) \ar^{\mu}[r] & S(R) \ar[r] & S(R/I) \ar[r] & 1}
\end{equation}
\begin{example}
 If one applies Loday's relativization to the functor $S(R) = \St(R)$ one recovers the usual relative Steinberg group $\St(R, I)$
  as defined in~\cite{Kbook}[Ch. III, Definition~5.7]. Notice that $\St(R, I)$ is not a subgroup of $\St(R)$ but rather its central extension.
\end{example}

For the rest of this subsection $G, P, S$ denote three group-valued functors on the category of commutative rings satisfying the following assumptions:
\begin{enumerate} [label=(A\arabic*)]
 \item \label{req:left-exact} $G$ is left exact.
 \item \label{req:subfunc} $P$ is a subfunctor of $G$ such that for all $R$ the group $P(R)$ is a perfect normal subgroup of $G(R)$.
 \item \label{req:uce} There is a natural transformation $\varphi \colon S \to P$ such that $ \varphi_R \colon S(R) \to P(R)$ is a universal central extension for all $R$.
 %This means that $\varphi_R$ is a certain surjective and that its kernel is contained in the center of $S(R)$.
 \item \label{req:coeq} For every ring $R$ and an ideal $I \trianglelefteq R$ the functor $S$ has the property that \[S(R \times_{R/I} R) \cong S(R) \times_{S(R/I)} S(R).\]
\end{enumerate}
\begin{example}
 The obvious triple of functors satisfying the above assumptions is as follows.
 Take the infinite general linear group functor $\mathrm{GL}(-)$ as $G$, its infinite elementary subfunctor $E(-)$ as $P$
  and the infinite Steinberg group functor $\St(-)$ as $S$. %TODO: Why these 4 propeties hold? What is a relative central extension.
\end{example}

@@

%TODO: Give the definition of $D_{R, I}$ and of the category of Pairs
\begin{prop}\label{characterization}
For $(G, P, R)$ as above the following statements are true.
\begin{enumerate}
 \item There are natural isomorphisms
  \[ \K_2(G, P, R) \cong \Ker(S(R) \to G(R)), \ \K_3(G, P, R) \cong \HH_3(S(R), \ZZ). \]
 \item The map $\mu \colon S(R, I) \to S(R)$ is the universal relative central extension of the map $S(R) \to S(R/I)$.
       The kernel $C(R, I)$ of $\mu$ is naturally isomorphic to the third relative homology group $\HH_3(S(R), S(R/I), \ZZ)$.
 \item There is a natural exact sequence:
 \[ \xymatrix{\K_3(G, P, R) \ar[r] & \K_3(G, P, R/I) \ar@{->>}[r] & C(R, I).} \]
 \end{enumerate}   
\end{prop}
\begin{proof}
 The first claim follows from~\ref{req:subfunc} and~\ref{req:uce} using the standard properties of the plus-construction, see \cite[\S~IV.1]{Kbook}   (cf. Exercises~1.8--1.9 ibid.). %provided one verifies that all the intermediate isomorphisms are natural in $R$.
